Manuscript

### 2025-02-27 Manuscript_MD_ScoreReLiability.qmd
init laptop change

From the 22 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 3 phenomena were coded
dichotomously, as responses were right or wrong (2 cases), or expressed
agreement vs disagreement dichotomously (1 case). Subsequently, the 22
phenomena need to be identified where the meta-analytic mean effect size
is statistically significant different from zero. This will be done as a
first step of this manuscript’s analysis procedure. Detailed information
concerning the different phenomena and how they were measured can be
found at \[osf-link\], while a brief summary can be found in Table 1.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version X.XX. To estimate meta-analytic mean ES and
heterogeneity, the REML-estimatore is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a significance level of .05 is used.

Additionally, estimates of score reliability are derived. While not
without its criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are evaluated by computing the simple mean in responses
across items for each individual. Computing the mean of several items
implicitly assumes tau-equivalence, as each item contributes equally to
the individual’s test score. Therefore, Cronbach’s Alpha is the better
choice to estimate score reliability, as it avoids this mismatch in
assumptions between how the score is computed and how score reliability
is estimated. Using these estimates of Cronbach’s Alpha, the individual
estimates of Cohen’s d are corrected, according to @eq-dc, including the
corrected standard errors (@eq-SEdc). Thereby, estimates of ES are
computed, which are corrected for imperfect reliability. Again, a
random-effects meta-analysis is run using metafor, generating an
estimate of heterogeneity in corrected ES. Different indicators of
heterogeneity are readily available, such as $I^2$, $H^2$ or the
coefficient of variation. However, the descriptions found in H&S and W&D
discuss the absolute amount of heterogeneity in terms of variance
($\tau^2$) or standard deviation ($\tau$). These parameters discuss the
variability of the standardized ES in the population in terms of
variance or standard deviation. Therefore, we also make use of absolute
heterogeneity here.

@@

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-Cis. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In @tbl-meanES, the estimates and
tests on the meta-analytic average ES, both corrected and uncorrected,
can be found.

@@

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability ρxx‘ and its heterogeneity. While it may not be
self-evident in the equation itself, it is easy to construct cases where
heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in
ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
In @fig-hs, the results of employing @eq-HS to compute the heterogeneity
(variance) in corrected ES is highlighted. All else held equal, only the
individual level of mean reliability \$ \mu\*{\*\\rho{xx‘}}\$ is varied
from .5 to .8. Mean corrected ES $\mu_{\delta}$ is held constant at 1,
its variance $\tau^2_{\delta}$ held constant at .04 and the variance of
score reliability $\tau^2_{\rho_{xx‘}}$ held constant at .02.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eg-HS

@@

In @fig-hs, the red line corresponds to the variance of corrected ES.

If everything else is held constant, the variance in uncorrected ES can
be a function only of the mean level of score reliability
$\mu_{rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in uncorrected ES does cross the red line, representing
the heterogeneity in corrected ES. Depending on the mean level of score
reliability, heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES.

However, @eq-hs is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
score reliability ρxx‘, can not be independent parameters by definition.
ES $\delta$ is the “true” standardized mean difference, unaffected by
sampling or measurement error. The ES $\delta_0$ is standardized using
the total standard deviation, which is inflated by measurement error.
Equation 1, demonstrating how Cohen’s d is computed demonstrates how the
mean difference is standardized in this case. As demonstrated further
above in @eg-dct, ES $\delta$ is standardized using the true standard
deviation in the sense of CTT. The fact that the true standard deviation
parameter $\sigma_T$ is found both in @eq-dct and @eq-rel demonstrates
that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
independent variables. Therefore, the basic premise of @eq-hs is not
fulfilled, as the assumption of variable independence could not be
fulfilled.

@@

As stated in @eq-dct, correcting ES for score reliability can be
described through a correction of the pooled standard deviation:
$\sigma_T = \sqrt{\rho_{xx’} \sigma^2_x}$. Since $\rho_{xx'}$ is always
between zero and one, in the case of imperfect score reliability,
$\sigma_T$ will always be smaller than total $\sigma_x$. Additionally,
if the heterogeneity in $\sigma_X$ was introduced purely by differences
in measurement precision – the error score variance $\sigma_E^2$, this
would be removed by the attenuation correction. However, alternatively,
the underlying latent variable we attempted to measure may not be
identically distributed across replications. In that case, true score
variance $\sigma^2_T$ would vary across replications, leading to
differences in both score reliability, corrected ES and raw ES.
Importantly, this implies that even if measurements of perfect score
reliability were taken across the different replications, heterogeneity
in ES would still persist.

In that case, some heterogeneity in $\sigma_T$ would remain, albeit
heterogeneity found in $\sigma_x$. Equation X essentially demonstrates
how mean and heterogeneity of MD, paired with mean and heterogeneity of
total score standard deviation $\sigma_x$ $\sigma_x$ can be used to
estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
Equation X to describe how, instead of total score standard deviation,
mean and variance in true score standard deviation $\sigma_T$ \sigma\_T
affect heterogeneity in corrected ES $\delta$.

@@

The differences between the two [Equations @eq-ratio_d0] and
[-@eq-ratio_d] demonstrates how ES heterogeneity is affected by an
attenuation correction procedure. According to the quotes in H&S and
W&D, @eq-ratio_d0 should lead to a larger value of
$\tau^2\left[\delta_0\right]$, compared to the estimate of
$\tau^2\left[\delta\right]$ from @eq-ratio_d. This, however, does not
follow from any equations shown throughout the text.

What is guaranteed is that the expected value of $\sigma_T$ is smaller
than the expected value of $\sigma_X$. Similarly, the variance in
$\sigma_T$ is guaranteed to be smaller than the uncorrected variance in
$\sigma_X$. However, since the expected value of either standard
deviation is placed in the denominator of the function and the variance
of either is placed in the numerator of the function, @eq-hs is not
sufficient to explain how heterogeneity in ES changes due to the
attenuation correction. Instead, we introduce two additional metrics.

@@

$R_1$ describes the relative size of mean true score variance, compared
to the mean observed score variance. This informs us, how much of the
mean total score variance can be attributed to mean true score variance.
Therefore, this metric is equivalent to the average score reliability.
Large values indicate that correcting the individual score variances (or
standard deviations for that matter) using the attenuation correction
procedure should lead to smaller change. Small values indicate the
opposite, large amounts of random error score variance on average lead
to large changes in mean score variance (or mean standard deviation).

@@

The metric $R_2$ on the other hand describes, in how far the attenuation
correction procedure has successfully reduced heterogeneity in score
variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
heterogeneity in true score variance $\tau^2_{\sigma^2_T}$, the score
variance after the attenuation correction, is essentially just as large
as the heterogeneity initally observed. Smaller values indicate how much
heterogeneity is "left", after applying an attenuation correction,
relative to the inital heterogeneity. This means that a value of .7
indicates that about 70% of heterogeneity in score variance remains,
even after correcting for differences in score reliability. As only 30%
of score variance heterogeneity could be removed, this would imply that
differences in error score variance $\sigma^2_E$ were responsible for
less than a third of the heterogeneity in score variances found. On the
contrary, more than two thirds of heterogeneity could be attributed to
actual differences in how the underlying true scores are distributed
across samples.

@@

Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
[-@eq-R2], we attempt to understand under which circumstances the
inequality defined in @eq-ineq_init holds true. However, while the
metrics make use of mean and heterogeneity of the score variances,
@eq-ineq_init makes use of mean and heterogeneity of the standard
deviations. In order to incorporate the metrics, it is necessary to
reparameterised the equation. Using the delta method to approximate mean
value and heterogeneity of variance from those of the standard
deviations, we know that (ref):
$$\mu_{\sigma^2_X} \approx \mu^2_{\sigma_X} \quad \quad  and \quad \quad 
CV_{\sigma^2} \approx 2 CV_{\sigma}$$ {#eq-delta_method}

@@

$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{1}{4} CV^2_{\sigma^2_t} \mu_{\sigma^2_t} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_CV}

@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
than the heterogeneity in uncorrected ES, using parameters of the
distributions of mean differences MD, true score variances\*
$\sigma^2_T$ and observed score variances $\sigma^2_X$. By introducing
the metrics $R_1$ and $R_2$ to this inequality, we can begin to
disentangle which circumstances need to be fulfilled for the inequality
to hold. Rearranging the terms in [Equations @eq-R1] and [-@eq-R2], and
subsequently entering these terms into Equation 18 leads to the
following
$$ \frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{1}{4} R^2_2 CV^2_{\sigma^2_X} R_1 \mu_{\sigma^2_x} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_R}

@@

Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
as both contain different, strictly positive parameters of the
distributions of true and observed score variance. Additionally, we know
that $R_1$ is equivalent to the average score reliability, and therefore
bound between $0\le R_1\le1$. Therefore, the term inside the first
bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similalry, $R_2$,
as the ratio of true and error score variance heterogeneity, is bound
between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
inside the second bracket turns negative. Generally, this means that the
inequality can only be violated if the reduction in relative
heterogeneity of score variance ($R_2$) is larger than the the average
score reliability to the power of 3.

Concerning the two scenarios, for scenario (a) to hold true, since the
term involving only $R_1$ in the first bracket can only be positive, the
second term involving both metrics would need to be negative. For the
first term to positively dominate the second term, however, we would
need to make additional assumptions about the size of the parameters
outside of the brackets, specifically about $\tau^2_{MD}$ and
$\mu_{MD}$. This is something we would prefer to avoid, as it involves
additional assumptions about the individual raw effect sizes of the
different phenomena. Additionally, it can be said that as long as the
term in the second bracket involving both metrics is positive, scenario
(b) is true by default. This occurs any time $R_1^3 > R_2$. For example,
if the average score reliability is about .8, if $R_2$ is at least .512
or larger, the inequality holds true and ES heterogeneity grows larger
as a consequence of the attenuation correction procedure, contrary to
claims made in W&D and H&S. This means that for the exemplary score
reliability of .8, a reduction in relative heterogeneity in score
variance of 49.8% ($1 – R_2$) or more is necessary for scenario (b) to
no longer be true. The smaller the average score reliability, the
smaller the $R_2$-metric may be for the inequality to still hold true.

Keeping this in mind, even if $R_2$ is smaller than .512, it does not
immediately imply that ES heterogeneity is smaller in $\delta$. For
example, it is unlikely that an $R_2 = .51$ leads to a term that is not
positively dominated by the first term involving only $R_1$. However,
where exactly the cut-off is, when ES heterogeneity starts shrinking due
to the attenuation correction procedure, can not be easily defined
without additional assumptions about the distribution of $MD$.

@@

Table 3 showed that for all phenomena, where we identified statistically
significant ES heterogeneity, the heterogeneity was even larger after
the attenuation correction procedure. Subsequently, in Equation 16 we
demonstrate that if the $R_2$ metric is larger than
$\sqrt{\mu[\rho_{xx’}]}$, ES heterogeneity is bound to be larger after
applying said procedure. Table 5 reports the meta-analytic estimates of
observed and true score variances, including their relative
heterogeneity and metrics $R_1$ and $R_2$ across all 12 phenomena. In
Table 5 we see that for ten out of 12 phenomena, the $R_2$ metric is
larger than 1. This implies that for all measurements concerning these
phenomena, the relative heterogeneity in observed score variances could
not be substantially explained by differences in measurement quality.
Correcting for heterogeneity in error score variance has in fact only
increased the extent of relative heterogeneity present in the score
variances. As demonstrated in Equation 20, this implies that the ES
heterogeneity grows even larger, as we correct for differences in score
reliability.

In @tbl-tauES we also found some phenomena where the heterogeneity did
in fact shrink after applying the attenuation correction procedure. This
is not reflected by what we observe in @tbl-R1R2, as $R_2 > R_1$ for
almost all phenomena, which would imply that across all phenomena we
expected ES heterogeneity to grow as a consequence of the attenuation
correction procedure. However, it is crucial to realise that we did not
find statistically significant heterogeneity for any of the phenomena
where ES heterogeneity was reduced. This implies, that we actually could
not statistically distinguish the ES heterogeneity from zero. Therefore,
what seems to be a reduction in ES heterogeneity was most likely caused
by estimation issues, as we could not accurately estimate the extent of
ES heterogeneity, either before or after the attenuation correction.

@@

Lastly, for two out of the 12 phenomena, we find a $CV_{\sigma^2_X}$ of
0, as the estimate of absolute heterogeneity $\tau_{\sigma^2_X}$ was 0
as well. For these phenomena, it is therefore also impossible to compute
the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
did not identify statistically significant ES heterogeneity in the first
place, we did find some for the Shnabel_Willingness_Reconcile_Rev
phenomenon. At the same time, we did not identify statistically
significant heterogeneity in score reliability. With zero heterogeneity
in score variances, Equation 13 implies that the ES heterogeneity is
bound to grow larger due to the attenuation procedure. Also, this
phenomenon was assessed with a particularly low number of replications
(8). Most likely, adequate power to detect heterogeneity in score
reliability and observed score variance requires a larger number of
replications than adequate power to detect heterogeneity. In that case,
the results from Table 5, concerning the
Shnabel_Willingness_Reconcile_Rev phenomenon might alternatively be
explained by power-issues. Similarly, while Table 3 does not imply
specific expectations concerning the presence of score variance
heterogeneity in the Alter_Analytic_Processing phenomenon, the low
number of replications (20) makes it hard to distinguish whether the
estimate of $\tau_{\sigma^2_X} = 0$ is actually sensible or can also be
attributed to low power.

@@

>>>>

From the 22 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 3 phenomena were coded
dichotomously, as responses were right or wrong (2 cases), or expressed
agreement vs disagreement dichotomously (1 case). Subsequently, those
phenomena need to be identified where the meta-analytic mean effect size
is statistically significant different from zero. This will be done as a
first step of this manuscript’s analysis procedure. Detailed information
concerning the different phenomena and how they were measured can be
found at \[osf-link\], while a brief summary can be found in Table A.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version X.XX. To estimate meta-analytic mean ES and
heterogeneity, the REML-estimator is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a significance level of .05 is used.

Additionally, estimates of score reliability are derived. While not
without criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are scored by computing the simple mean or sum of
responses across items for each individual. Computing the mean or sum of
several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. Using these estimates of
Cronbach’s Alpha, the individual estimates of Cohen’s d are corrected,
according to @eq-dc, including the corrected standard errors (@eq-SEdc).
Thereby, estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation. However, the descriptions found in H&S
and W&D discuss the absolute amount of heterogeneity in terms of
variance ($\tau^2$) or standard deviation ($\tau$). These parameters
discuss the variability of the standardized ES in the population in
terms of variance or standard deviation. Therefore, we also make use of
absolute heterogeneity here.

@@

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-CIs. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In @tbl-meanES, the estimates and
tests on the meta-analytic average ES, both corrected and uncorrected,
can be found.

@@

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability $\rho_{xx‘}$ and its heterogeneity. While it may
not be self-evident in the equation itself, it is easy to construct
cases where heterogeneity in $\delta_0$ is inflated or deflated by
heterogeneity in $\rho_{xx‘}$, contradicting Hunter & Schmidt’s
interpretation of this equation. In @fig-HS, the results of employing
@eq-HS to compute the heterogeneity (variance) in corrected ES is
highlighted. All else held equal, only the individual level of mean
reliability $\mu_{\rho_{xx‘}}$ is varied from .5 to .8. Mean corrected
ES $\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
held constant at .04 and the variance of score reliability
$\tau^2_{\rho_{xx‘}}$ held constant at .02.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eq-HS

@@

In @fig-HS, the dashed red line corresponds to the variance of corrected
ES.

If everything else is held constant, the variance in uncorrected ES can
be a function only of the mean level of score reliability
$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in uncorrected ES does cross the red line, representing
the heterogeneity in corrected ES. This implies that, as long as mean
score reliability is below .72, an attenuation correction would lead to
a reduction in ES heterogeneity. As mean score reliability grows larger
than .72, however, the heterogeneity in uncorrected ES grows larger than
the heterogeneity in corrected ES - an attenuation correction would
therefore only increase the extent of ES heterogeneity. This
demonstrates that, depending on the mean level of score reliability,
heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES. As heterogeneity in corrected ES is not
always smaller, the interpretationsof H&S concerning this equation is
erroneous.

However, @eq-HS is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, @eq-HS is only valid “If the
level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
score reliability ρxx‘, can not be independent parameters by definition.
ES $\delta$ is the “true” standardized mean difference, unaffected by
sampling or measurement error. The ES $\delta_0$ is standardized using
the total standard deviation, which is inflated by measurement error.
Equation 1, demonstrating how Cohen’s d is computed demonstrates how the
mean difference is standardized in this case. As demonstrated further
above in @eg-dct, ES $\delta$ is standardized using the true standard
deviation in the sense of CTT. The fact that the true standard deviation
parameter $\sigma_T$ is found both in @eq-dct and @eq-rel demonstrates
that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
independent variables. Therefore, the basic premise of @eq-HS is not
fulfilled, as the assumption of variable independence could not be
fulfilled.

@@

As stated in @eq-dct, correcting ES for score reliability can be
described through a correction of the pooled standard deviation:
$\sigma_T = \sqrt{\rho_{xx’} \sigma^2_x}$. Since $\rho_{xx'}$ is always
between zero and one, in the case of imperfect score reliability,
$\sigma_T$ will always be smaller than total $\sigma_x$. Additionally,
if the heterogeneity in $\sigma_X$ was introduced purely by differences
in measurement precision – the error score variance $\sigma_E^2$, this
would be removed by the attenuation correction. However, alternatively,
the underlying latent variable we attempted to measure may not be
distributed identically across replications. In that case, true score
variance $\sigma^2_T$ would vary across replications, leading to
differences in both score reliability, corrected ES and raw ES.
Importantly, this implies that even if measurements of perfect score
reliability were taken across the different replications, heterogeneity
in ES would still persist.

In that case, some heterogeneity in $\sigma_T$ would remain, albeit
heterogeneity found in $\sigma_x$ was reduced. @eq-ratio_d0 essentially
demonstrates how mean and heterogeneity of MD, paired with mean and
heterogeneity of total score standard deviation $\sigma_x$ can be used
to estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
Equation @eq-ratio_d0 to describe how, instead of total score standard
deviation, mean and variance in true score standard deviation $\sigma_T$
affect heterogeneity in corrected ES $\delta$.

@@

The differences between the two [Equations @eq-ratio_d0] and
[-@eq-ratio_d] demonstrates how ES heterogeneity is affected by an
attenuation correction procedure. According to the quotes in H&S and
W&D, @eq-ratio_d0 should lead to a larger value of
$\tau^2\left[\delta_0\right]$, compared to the estimate of
$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, the equations
shown throughout this text do not agree with this conclusion.

Instead, what is guaranteed is that the expected value of $\sigma_T$ is
smaller than the expected value of $\sigma_X$. Similarly, the
heterogeneity in $\sigma_T$ is guaranteed to be smaller than the
uncorrected heterogeneity in $\sigma_X$. However, since the expected
value of either standard deviation is placed in the denominator of the
function and the variance of either is placed in the numerator of the
function, @eq-HS is not sufficient to explain how heterogeneity in ES
changes due to the attenuation correction. Instead, we introduce two
additional metrics.

@@

$R_1$ describes the relative size of mean true score variance, compared
to the mean observed score variance. This informs us, how much of the
mean total score variance can be attributed to mean true score variance.
Therefore, this metric is close to the average score reliability. Even
though formally $R_1$ and the meta-analytic mean score reliability
$\mu_{\rho_{xx'}}$ are not the same value, in practice we observe little
to no differences between the two parameters. Large values indicate that
correcting the individual score variances (or standard deviations for
that matter) using the attenuation correction procedure should lead to
very little change, as most of the variation observed were due to
genuine differences in true scores. Small values indicate the opposite,
the observed score variance was due to a larger amount of random error
score variance, leading to a stronger correction.

@@

The metric $R_2$ on the other hand describes, in how far the attenuation
correction procedure has successfully reduced heterogeneity in score
variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
heterogeneity in true score variance $\tau^2_{\sigma^2_T}$, the score
variance after the attenuation correction, is essentially just as large
as the heterogeneity initially observed. Smaller values indicate how
much heterogeneity is "left", after applying an attenuation correction,
relative to the initial heterogeneity. This means that a value of .7
indicates that about 70% of heterogeneity in score variance remains,
even after correcting for differences in score reliability. As only 30%
of score variance heterogeneity could be removed, this would imply that
differences in error score variance $\sigma^2_E$ were responsible for
less than a third of the heterogeneity in score variances found. On the
contrary, more than two thirds of heterogeneity could be attributed to
actual differences in how the underlying true scores are distributed
across samples.

Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
of true and total score variance, not standard deviations. This choice
was made to facilitate the role of score reliability in the subsequent
discussion, as score reliability itself is defined by variances, not
standard deviations (see @eq-rel).

@@

Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
[-@eq-R2], we attempt to understand under which circumstances the
inequality defined in @eq-ineq_init holds true. However, while the
metrics make use of mean and heterogeneity of the score variances,
@eq-ineq_init makes use of mean and heterogeneity of the standard
deviations. In order to incorporate the metrics, it is necessary to
reparameterise the equation. Using the delta method to approximate mean
value and heterogeneity of variance from those of the standard
deviations, we know that (ref):
$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}

@@

$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}

@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
than the heterogeneity in uncorrected ES, using parameters of the
distributions of mean difference MD, true score variance $\sigma^2_T$
and observed score variance $\sigma^2_X$. By introducing the metrics
$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
conditions need to be fulfilled for the inequality to hold. Rearranging
the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
these terms into Equation 18 leads to the following
$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}

@@

Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
as both contain different, strictly positive parameters of the
distributions of true and observed score variance. Additionally, we know
that $R_1$ is equivalent to the average score reliability, and therefore
bound between $0\le R_1\le1$. Therefore, the term inside the first
bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
as the ratio of true and error score variance heterogeneity, is bound
between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
inside the second bracket turns negative. Generally, this means that the
inequality can only be violated if the reduction in relative
heterogeneity of score variance ($R_2$) is larger than the the average
score reliability to the power of 3.

Concerning the two scenarios, for scenario (a) to hold true, since the
term involving only $R_1$ in the first bracket can only be positive, the
second term involving both metrics would need to be negative. For the
first term to positively dominate the second term, however, we would
need to make additional assumptions about the size of the parameters
outside of the brackets, specifically about $\tau^2_{MD}$ and
$\mu_{MD}$. This is something we would prefer to avoid, as it involves
additional assumptions about the individual raw effect sizes of the
different phenomena. Additionally, it can be said that as long as the
term in the second bracket involving both metrics is positive, scenario
(b) is true by default. This occurs any time $R_1^3 > R_2$. For example,
assuming the average score reliability is about .8, if $R_2$ is at least
.512 or larger, the inequality holds true and ES heterogeneity grows
larger as a consequence of the attenuation correction procedure,
contrary to claims made in W&D and H&S. This means that for the
exemplary score reliability of .8, a reduction in relative heterogeneity
in score variance of 48.8% ($(1 – R_2)100\%$) or more is necessary for
scenario (b) to no longer be true. The smaller the average score
reliability, the smaller the $R_2$-metric may be for the inequality to
still hold true.

Keeping this in mind, even if $R_2$ is smaller than .512, it does not
immediately imply that ES heterogeneity is smaller in $\delta$. For
example, it is unlikely that an $R_2 = .5$ leads to a term that is not
positively dominated by the first term involving only $R_1$. However,
where exactly the cut-off is, where ES heterogeneity starts shrinking
due to the attenuation correction procedure, can not be easily defined
without additional assumptions about the distribution of $MD$.

@@

@tbl-tauES showed that for all phenomena, where we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Subsequently, in
Equation 16 we demonstrate that if the $R_2$ metric is larger than
$R_1^3$, ES heterogeneity is bound to be larger after applying said
procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
true score variances, including their heterogeneity and metrics $R_1$
and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across nine phenomena, $R_2$ is larger
than .97. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution. Next to the two exceptional phenomena,
where zero heterogeneity in score variance was found
(Alter_Analytic_Processing and Shnabel_Willingness_Reconcile_Rev), only
for the Srull_Ronald_Hostility phenomenon some heterogeneity could be
attributed to differences in random error variance, albeit less than
14%.

This implies that for all measurements concerning these phenomena, the
heterogeneity in observed score variances could not be substantially
explained by differences in measurement quality. For no phenomenon, if
differences in score variance could be found, was $R_2$ smaller than
$R_1^3$. As demonstrated in Equation 20, this implies that the ES
heterogeneity grows even larger, as we correct for differences in score
reliability.

@@

In @tbl-tauES we also found some phenomena where the heterogeneity did
in fact shrink after applying the attenuation correction procedure. This
is not reflected by what we observe in @tbl-R1R2, as $R_2 > R_1$ for
almost all phenomena, which would imply that across all phenomena we
expected ES heterogeneity to grow as a consequence of the attenuation
correction procedure. However, it is crucial to realise that we did not
find statistically significant ES heterogeneity for any of the phenomena
where ES heterogeneity was reduced. This implies, that we actually could
not statistically distinguish the ES heterogeneity from zero. Therefore,
what seems to be a reduction in ES heterogeneity was most likely caused
by estimation issues, as we could not accurately estimate the extent of
ES heterogeneity, either before or after the attenuation correction.

Lastly, for two out of the 12 phenomena, we find a $\tau_{\sigma^2_X}$
of 0. For these phenomena, it is therefore also impossible to compute
the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
did not identify statistically significant ES heterogeneity in the first
place, we did find some for the Shnabel_Willingness_Reconcile_Rev
phenomenon. At the same time, we did not identify statistically
significant heterogeneity in score reliability. With zero heterogeneity
in score variances, Equation 13 implies that the ES heterogeneity is
bound to grow larger due to the attenuation procedure. Also, this
phenomenon was assessed with a particularly low number of replications
(8). Most likely, adequate power to detect heterogeneity in score
reliability and observed score variance requires a larger number of
replications than adequate power to detect heterogeneity. In that case,
the results from Table 5, concerning the
Shnabel_Willingness_Reconcile_Rev phenomenon might alternatively be
explained by power-issues. Similarly, while Table 3 does not imply
specific expectations concerning the presence of score variance
heterogeneity in the Alter_Analytic_Processing phenomenon, the low
number of replications (20) makes it hard to distinguish whether the
estimate of $\tau_{\sigma^2_X} = 0$ is actually sensible or can also be
attributed to low power.

@@

The data-sets available to this project fall into two categories: direct
multi-site replications of a specific phenomenon, employing identical
protocols across samples; a large-scale collaborative effort to assess a
single phenomenon globally, distributed across a large number of
independent laboratories. For both categories, rather arbitrary sampling
schemes were used. The majority of respondents are students at the
respective universities, taking part in the studies voluntarily or
because they were required as part of their study's programme (at times
compensation was awarded). This is not the same as random sampling.
These samples are put together arbitrarily, it is unlikely that the
underlying distribution of true scores is meaningful for a properly
specified population. The between-studies randomness introduced by these
undefined sampling schemes manifests itself in true score variance
heterogeneity $\tau^2_{\sigma^2_X}$. Between-studies randomness
introduces by issues of measurement quality manifest themselves in error
score variance heterogeneity $\tau^2_{\sigma^2_X }$. This implies that
it should be of little surprise that the differences in true score
variance largely exceed those in error score variance.
